---
title: 'Brandi Christiano - Homework #3 Notebook'
author: "BMC"
date: "4/21/2021"
output: html_document
---

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(tidyverse)
library(tidyr)
library (ggpmisc)
library(dplyr)
library(broom)
```

```{r setup, include=FALSE}
df <- read.csv("KamilarAndCooperData.csv")
```

```{r}
head(df)
```

# **Challenge 1**

## Normal Model (Model1)

### Fit the regression model 
```{r}
# Normal Model
model1 <- lm(WeaningAge_d~Brain_Size_Species_Mean, data=df)
model1
```
###  Using {ggplot2}, produce a scatterplot with the fitted line superimposed upon the data
```{r, warning=FALSE}

# x is the control variable
plot1 <- ggplot(data = df, aes(x =(Brain_Size_Species_Mean), y =(WeaningAge_d))) + geom_point() + geom_smooth(method = "lm", formula = y ~ x) 
plot1
```


### Append the the fitted model equation to your plot.
```{r, warning=FALSE}
# Make the output of Model 1 an Equation to include on the plot 

eq <- substitute(italic(y) == a + b *"x" , 
         list(a = format(unname(coef(model1)[1]), digits = 2),
              b = format(unname(coef(model1)[2]), digits = 2)))
eq <- c("y == 132+2.6(x)")
eq

plot1 <- plot1+ geom_text(x = 400, y = 1500, label =eq, parse = TRUE)

plot1
```

### Identify and interpret the point estimate of the slope (β1)
 
```{r}
# Model 1 Betas 
model.summary <- tidy(model1)

beta0 <- model.summary %>%
  filter(term == "(Intercept)") %>%
  pull(estimate)

beta1 <- model.summary %>%
  filter(term == "Brain_Size_Species_Mean") %>%
  pull(estimate)

beta0 # 132.46
beta1 # 2.64

# INTERPERTATION: Beta 1 is the true slope, therefore every time the  x variable (Mean of Brain Size) increases by 1, the y variable (Weaning Age in Days), will increase by Beta 1, which is 2.64. 

# Test null hypothesis that beta1 = 0
beta1==0 # False

# Test alternative hypothesis that beta1 does not = 0
beta1!=0 # True

```
### Find a 90% CI for the slope (β1) parameter
```{r}
# Model 1 CIs
alpha <- 0.25
ci <- predict(model1,
  newdata = data.frame(Brain_Size_Species_Mean = 2.64),
  interval = "confidence", level = 1 - alpha
) # for a single value
ci

# The confidence intervals range from 117.6 - 161.2
```
### Using your model, add lines for the 90% confidence
```{r, warning=FALSE}
# Model 1

df2 <- augment(model1, se_fit = TRUE)
head(df2)
df2
df$Brain_Size_Species_Mean

d <- mutate(df, centered_Age = WeaningAge_d - mean(WeaningAge_d))
d <- mutate(df, centered_Brain = Brain_Size_Species_Mean - mean(Brain_Size_Species_Mean))

d$Brain_Size_Species_Mean

newdata <-  data.frame(Brain_Size_Species_Mean = d$Brain_Size_Species_Mean)
newdata

ci <- predict(model1,newdata,
              interval = "confidence", level = 1 - alpha) 

ci
ci <- data.frame(ci)
ci
ci <- cbind(df$Brain_Size_Species_Mean, ci)
ci
names(ci) <- c("Brain.Size.Species.Mean", "c.fit", "c.lwr", "c.upr")
ci
g <- ggplot(data = df, aes(x = Brain_Size_Species_Mean, y = WeaningAge_d))
g
g <- g + geom_point(alpha = 0.1)
g
g2 <- g + geom_line(
  data = ci, aes(x =ci$Brain.Size.Species.Mean, y = c.fit),
  color = "black"
)
g2
g3 <- g2 + geom_line(
  data = ci, aes(x =ci$Brain.Size.Species.Mean , y = c.lwr),
  color = "blue"
)
g3
g4 <- g3 + geom_line(
  data = ci, aes(x = ci$Brain.Size.Species.Mean, y = c.upr),
  color = "blue"
)
g4

```



### Using your model, add lines for prediction interval bands on the plot
```{r, warning=FALSE}
pi <- predict(model1, newdata ,
  interval = "prediction", level = 0.90
) # for a vector of values
pi <- data.frame(pi)
pi
pi <- cbind(df$Brain_Size_Species_Mean, pi)
names(pi) <- c("Brain.Size.Species.Mean", "p.fit", "p.lwr", "p.upr")
pi
g5 <- g4 + geom_line(data = pi, aes(x = Brain.Size.Species.Mean, y = p.lwr), color = "red")
g5
g6 <- g5 + geom_line(data = pi, aes(x = Brain.Size.Species.Mean, y = p.upr), color = "red")
g6 



```



### Add a legend to differentiate between the lines
```{r}
# Normal Model
```

### Produce a point estimate for the weaning age of a species whose brain weight is 750 mg. 

```{r}
# Normal Model
beta0 <- model.summary %>%
  filter(term == "(Intercept)") %>%
  pull(estimate)
beta1 <- model.summary %>%
  filter(term == "Brain_Size_Species_Mean") %>%
  pull(estimate)
(h.hat <- beta1 * 750 + beta0)

# 2110.3
```
### Produce the  associated 90% prediction interval for the weaning age of a species whose brain weight is 750 mg.
```{r}
# Normal Model
pi <- predict(model1,
  newdata = data.frame(Brain_Size_Species_Mean = 750),
  interval = "prediction", level = 0.90
) # for a single value
pi
# 1783.8 - 2436.8
```






















## Log Model (Model2)

### Fit the regression model 
```{r, warning=FALSE}
# Log Model
model2 <- lm(log(WeaningAge_d)~log(Brain_Size_Species_Mean), data=df)
model2

```
###  Using {ggplot2}, produce a scatterplot with the fitted line superimposed upon the data
```{r, warning=FALSE}
# Log Model Plot 
plot2 <- ggplot(data = df, aes(x = log(Brain_Size_Species_Mean), y = log(WeaningAge_d))) + geom_point() + geom_smooth(method = "lm", formula = y ~ x) 
```
### Append the the fitted model equation to your plot.
```{r, warning=FALSE}
eq2 <- substitute(italic(y) == a + b *"x" , 
         list(a = format(unname(coef(model2)[1]), digits = 2),
              b = format(unname(coef(model2)[2]), digits = 2)))
eq2 <- c("y == 3.4+0.56(x)")
eq2

plot2 <- plot2+ geom_text(x = 2, y = 6.5, label =eq2, parse = TRUE)

plot2
```
### Identify and interpret the point estimate of the slope (β1)
```{r}
# Log Model
model.summary2 <- tidy(model2)

beta0 <- model.summary2 %>%
  filter(term == "(Intercept)") %>%
  pull(estimate)

beta1 <- model.summary2 %>%
  filter(term == "log(Brain_Size_Species_Mean)") %>%
  pull(estimate)

beta0 # 3.37
beta1 # 0.57

# INTERPERTATION: Beta 1 is the true slope, therefore every time the  x variable (the log of the mean of brain size) increases by 1, the y variable (the log of weaning age in days), will increase by Beta 1, which is 0.57. 

# Test null hypothesis that beta1 = 0
beta1==0 # False

# Test alternative hypothesis that beta1 does not = 0
beta1!=0 # True
```
###  Find a 90% CI for the slope (β1) parameter.
```{r}
# Model 2 CIs
alpha <- 0.25
ci2 <- predict(model2,
  newdata = data.frame(Brain_Size_Species_Mean = 0.57),
  interval = "confidence", level = 1 - alpha
) # for a single value
ci2

# The confidence intervals range from 2.9 - 3.2
```
```{r, warning=FALSE}
# Log Model

df3 <- augment(model2, se_fit = TRUE)
head(df3)
df3
df$Brain_Size_Species_Mean

d <- mutate(df, centered_Age = log(WeaningAge_d) - mean(log(WeaningAge_d)))
d <- mutate(df, centered_Brain = log(Brain_Size_Species_Mean) - mean(log(Brain_Size_Species_Mean)))
d


newdata <-  data.frame(Brain_Size_Species_Mean = d$Brain_Size_Species_Mean)
newdata

ci <- predict(model2,newdata,
              interval = "confidence", level = 1 - alpha) 

ci
ci <- data.frame(ci)
ci


BrainLog <- log(df$Brain_Size_Species_Mean)

ci <- cbind(BrainLog, ci)
ci
names(ci) <- c("Log.Brain.Size.Species.Mean", "c.fit", "c.lwr", "c.upr")
ci
g <- ggplot(data = df, aes(x = log(Brain_Size_Species_Mean), y = log(WeaningAge_d)))
g
g <- g + geom_point(alpha = 0.1)
g
g2 <- g + geom_line(
  data = ci, aes(x =ci$Log.Brain.Size.Species.Mean, y = c.fit),
  color = "black"
)
g2
g3 <- g2 + geom_line(
  data = ci, 
  aes(x =ci$Log.Brain.Size.Species.Mean, y = c.lwr),
  color = "blue"
)
g3
g4 <- g3 + geom_line(
  data = ci, aes(x = ci$Log.Brain.Size.Species.Mean, y = c.upr),
  color = "blue"
)
g4
```
### Using your model, add lines for prediction interval bands on the plot

```{r, warning=FALSE}
pi <- predict(model2, newdata ,
  interval = "prediction", level = 0.90
) # for a vector of values
pi <- data.frame(pi)
pi
pi <- cbind(BrainLog, pi)
names(pi) <- c("Log.Brain.Size.Species.Mean", "p.fit", "p.lwr", "p.upr")
pi
g5 <- g4 + geom_line(data = pi, aes(x = Log.Brain.Size.Species.Mean, y = p.lwr), color = "red")
g5
g6 <- g5 + geom_line(data = pi, aes(x = Log.Brain.Size.Species.Mean, y = p.upr), color = "red")
g6 

```



### Produce a point estimate for the weaning age of a species whose brain weight is 750 gm.
```{r}
# Log Model
log(750) # = 6.62
beta0 <- model.summary2 %>%
  filter(term == "(Intercept)") %>%
  pull(estimate)
beta0
beta1 <- model.summary2 %>%
  filter(term == "log(Brain_Size_Species_Mean)") %>%
  pull(estimate)
beta1
(h.hat <- beta1 * 6.62 + beta0)
# 7.2
```

### Produce the  associated 90% prediction interval for the weaning age of a species whose brain weight is 750 gm.
```{r}
#Log Model

pi2 <- predict(model2,
  newdata = data.frame(Brain_Size_Species_Mean = 6.62),
  interval = "prediction", level = 0.90
) # for a single value
pi2
# 3.8 - 5.1
```

###  Q - Do you trust the model to predict observations accurately for this value of the explanatory variable? Why or why not?
### A -  No, I think the variables are too skewed for the normal model to predict an accurate observation. I would trust the log transformations more. 

### Q - Looking at your two models (i.e., untransformed versus log-log transformed), which do you think is better? Why?
### A -  Because the variables in the untransformed model are highly skewed, I think the log-transformed model would be better. The log-transformation creates a  more normalized data set and results in a more linear relationship. 





# **Challenge 2**
